---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---



## Loading and preprocessing the data
```{r, echo=TRUE}
activity = read.csv('activity.csv')
str(activity)
```
```{r}
summary(activity)
```

```{r}
head(activity)
```

```{r}
tail(activity)
```


## What is mean total number of steps taken per day?

```{r}
dailySteps <- aggregate(steps ~ date, data = activity, FUN=sum,na.rm=T)
theMean <- mean(dailySteps$steps,na.rm=T)
theMedian <- median(dailySteps$steps,na.rm=T)
hist(dailySteps$steps, xlab="steps",ylab="Frequency",
     main="Steps per day distribution",border='orange',col="gray50")
abline(v = theMean, col = "blue", lwd = 5)
abline(v = theMedian, col = "red", lwd = 2)
```


```{r}
summary(dailySteps)
```

So Mean is `r theMean` and Median is `r theMedian`


## What is the average daily activity pattern?

```{r}
fiveMinSteps <- aggregate(steps ~ interval, data = activity, FUN=mean, na.rm=T)
plot(steps~interval,data=fiveMinSteps,type="l")
```

#Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxInterval <- fiveMinSteps[which.max(fiveMinSteps$steps),]
maxInterval
```

**`r maxInterval$interval`'s**  5-minute interval, on average across all the days in the dataset, contains the maximum number of steps. And steps values is **`r maxInterval$steps`**

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
