---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---



## Loading and preprocessing the data
```{r, echo=TRUE}
activity = read.csv('activity.csv')
str(activity)
```

#### Overview
```{r}
summary(activity)
```

```{r}
head(activity)
```

```{r}
tail(activity)
```

#### Interval ID pattern

If we look at interval ID, we can find pattern how ID depends on time of the day in 24 hour format.
```{r}
unique(activity$interval)
```

For example, 0 - is 5 minutes intrval from 00:00 to 00:05.  
2355 - is the last 5 minutes interval from 23:55 to 00:00 next day.


## What is mean total number of steps taken per day?

```{r}
dailySteps <- aggregate(steps ~ date, data = activity, FUN=sum,na.rm=T)
theMean <- mean(dailySteps$steps,na.rm=T)
theMedian <- median(dailySteps$steps,na.rm=T)
hist(dailySteps$steps, xlab="steps",ylab="Frequency",
     main="Steps per day distribution",border='orange',col="gray50")
abline(v = theMean, col = "blue", lwd = 5)
abline(v = theMedian, col = "red", lwd = 2)
legend(19000, 25, c('mean','median'), lwd=4, col=c("blue", "red"))

```


```{r}
summary(dailySteps)
```

So Mean is **`r format(theMean, digits=2, nsmall=2)`** and Median is **`r theMedian`**


## What is the average daily activity pattern?

#### What is average number of steps taken, averaged across all days?

```{r}
fiveMinSteps <- aggregate(steps ~ interval, data = activity, FUN=mean, na.rm=T)
plot(steps~interval,data=fiveMinSteps,type="l")
```

#### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxInterval <- fiveMinSteps[which.max(fiveMinSteps$steps),]
maxInterval
```

**`r maxInterval$interval`'s**  5-minute interval, on average across all the days in the dataset, contains the maximum number of steps. And steps values is **`r maxInterval$steps`**

Which means that maximum average activity was **at 08:35 - 08:40 in the morning**.

## Imputing missing values

####  Calculate total number of NA's 

There are several ways to do it:

  * summary function 
```{r}
  summary(activity)
```
  * sum function
```{r}
  sum(is.na(activity$steps))
```
In both cases Total NA's number is 2304.


####  Create new Dataset with NA's filled in.

```{r}
df <- activity
df$steps[is.na(df$steps)] <- aggregate(steps ~ interval, data = df, FUN=mean,na.rm=T)$steps
summary(df)
```

There is no NA's compared to initial dataset and max value is Daily mean.

Initial dataset:
```{r}
summary(activity)
```


####  Build new histogram and compare with initial
```{r}
dailyStepsNoNA <- aggregate(steps ~ date, data = df, FUN=sum,na.rm=T)
theMeanNoNA <- mean(dailyStepsNoNA$steps,na.rm=T)
theMedianNoNA <- median(dailyStepsNoNA$steps,na.rm=T)
hist(dailyStepsNoNA$steps, xlab="steps",ylab="Frequency",
     main="Steps per day distribution for No NA Dataset",border='green',col="grey80")

abline(v = theMeanNoNA, col = "blue", lwd = 5)
abline(v = theMedianNoNA, col = "red", lwd = 2)
legend(19000, 35, c('mean','median'), lwd=4, col=c("blue", "red"))
```


#### Do these values differ from the estimates from the first part of the assignment? 
```{r}
summary(dailyStepsNoNA)
```

#### What is the impact of imputing missing data on the estimates of the total daily number of steps?

Number of days with mean steps increased upto 35.

Mean is **`r format(theMeanNoNA, digits=2, nsmall=2)`** and Median is **`r format(theMedianNoNA, digits=2, nsmall=2)`**. So they became equal.


## Are there differences in activity patterns between weekdays and weekends?

#### Add new factor to dataset

factor wd - lables if day is weekday or weekend

```{r}
df$dayType <- factor((weekdays(as.Date(df$date)) %in% c('Saturday','Sunday')), levels=c(TRUE,FALSE), labels=c('weekend', 'weekday'))
df$weekday <- weekdays(as.Date(df$date))
head(df)
```

#### Plot time series 
```{r}
fiveMinStepsWd <- aggregate(steps ~ interval + dayType, data = df, FUN=mean, na.rm=T)
library(ggplot2)
ggplot(fiveMinStepsWd, aes(interval, steps)) + 
        geom_line(color="blue")+ 
        facet_grid(dayType ~ .) +
        xlab("5-minute interval") + 
        ylab("avarage number of steps")
```